---
title: "covariables"
output: html_document
date: "2025-04-25"
format:
  html:
    toc: true
    toc-location: left
    theme: cerulean
    css: style.css
    embed-resources: true
    number-sections: true
---

```{r}
#| echo: false
source("_common.R")
knitr::opts_chunk$set(
  eval = TRUE,
  echo = FALSE
)

```

```{r }
#| include: false
# load libraries and functions-----------------------------------------------
source(file.path("..", "scripts", "Librerias.R"))
source(file.path("..", "scripts", "Funciones.R"))
conflicted::conflicts_prefer(purrr::compose)
library(fuzzyjoin)

```

```{css, echo = FALSE}
.output {
max-height: 500px;
overflow-y: scroll;
}

```

# Load data

```{r}
# load pathes----------------------------------------------------------------
brutos      <- file.path( "r:", "PREGVAL", "1-DATOS", 
                          "1-BRUTOS")
intermedios <- file.path( "r:", "PREGVAL", "1-DATOS", 
                          "2-INTERMEDIOS")
procesados  <- file.path( "r:", "PREGVAL", "1-DATOS", 
                          "3-PROCESADOS")

```

```{r}
# load pregnancies-----------------------------------------------------------
pregnancies_cohort_sip <- readRDS(file.path(procesados, "pregnancy_algorithm",
                                      "pregnancies_cohort_sip.RDS")) 

sips_pregnancies_cohort <- readRDS(file.path(procesados, "sips",
                                         "sips_pregnancies_cohort.RDS"))

```

# Study period

```{r}
# check population-----------------------------------------------------------
sips_all <- readRDS(file.path(procesados, "sips", "sips_all.RDS"))
sips_clean <- readRDS(file.path(procesados, "sips", "sips_clean.RDS"))

# filter study period--------------------------------------------------------
pregnancies_cohort_00 <- pregnancies_cohort_sip |> 
  # filter from sip
  tidylog::filter(sip %in% sips_all) |> 
  # filter from sip
  tidylog::filter(sip %in% sips_clean) 
  
```

# Risk of exclusion and birthdate

```{r}
# obtain exclusion_risk and birthdate----------------------------------------
pregnancies_cohort_0 <- pregnancies_cohort_sip |> 
  mutate(year = year(fecha_inicio)) |> 
  mutate(exclusion_risk = case_when(
    exclusion_risk_0 == 0 ~ "not at risk",
    exclusion_risk_0 == 1 ~ "at risk",
    exclusion_risk_0 == 2 ~ "at risk",
    exclusion_risk_0 == 3 ~ "at risk",
    exclusion_risk_0 == 4 ~ "at risk",
    exclusion_risk_0 == 5 ~ "at risk",
    exclusion_risk_0 == 6 ~ "at risk",
    exclusion_risk_0 == 7 ~ "at risk",
    is.na(exclusion_risk_0) ~ "no information available",
    T ~ "no information available") |> factor(levels = c(
      "not at risk", "at risk", "no information available")))

```

## Age

```{r}
# add age--------------------------------------------------------------------
pregnancies_cohort_0b <- pregnancies_cohort_0 |> 
  mutate(age = as.numeric(difftime(fecha_inicio, fecha_nacimiento)/365.25))

pregnancies_cohort_0b |> tidylog::filter(is.na(fecha_nacimiento))

pregnancies_cohort_0b |> summarise(
  mean_age = mean(age, na.rm = TRUE),
  sd_age = sd(age, na.rm = TRUE)
  )

# create cuts of age <=25, 26-35, > 35
pregnancies_cohort_0b |> 
  mutate(age_group = case_when(
    age < 26 ~ "<= 25",
    age >= 26 & age < 36 ~ "26-35",
    age >= 36 ~ "> 35",
    T ~ "<= 25") |> factor(levels = c("<= 25", "26-35", "> 35"))) |>  
  tabyl(age_group)

```


## pais de origen

```{r}
# country--------------------------------------------------------------------
pregnancies_cohort_1 <- pregnancies_cohort_0b |> 
    mutate(cod_country = str_sub(pais_nacimiento, 1, 3)) |>  
  mutate(country = case_when(
           cod_country == "108" ~ "Spain",
           cod_country != "108" & cod_country >= "102" & 
             cod_country < "133" ~ "Europe",
           cod_country >= "135" & cod_country < "999" ~ "Other",
           T ~ "Unknown")) |> 
  mutate(country = factor(country, 
                   levels = c("Spain", "Europe", "Other", "Unknown")))

pregnancies_cohort_1 |> tabyl(country) 

```

## Income

```{r}
# Income---------------------------------------------------------------------
# check rafs_primer
pregnancies_cohort_1 |> tabyl(raf_ipago_primer, raf_ilimi_primer)

# obtain exclusion_risk table
pregnancies_cohort_1b <- pregnancies_cohort_1 |> 
  mutate(raf_ipago2 = if_else(raf_ipago == "", 
                             raf_ipago_primer, raf_ipago),
         raf_ilimi2 = if_else(is.na(raf_ilimi),
                             raf_ilimi_primer, raf_ilimi))

pregnancies_cohort_1b |> count(raf_ilimi2)
pregnancies_cohort_1b |> count(raf_ipago2)

pregnancies_cohort_1b |> 
  mutate(income = case_when(
    raf_ipago2 == "TSI 000" ~ "Unknown",
    raf_ipago2 == "TSI 001" ~ "No income", 
    raf_ipago2 == "TSI 002" & raf_ilimi2 == "0" ~ "No income", 
    raf_ipago2 == "TSI 002" & raf_ilimi2 == "1" ~ "<18,000",
    raf_ipago2 == "TSI 002" & raf_ilimi2 == "2" ~ "18,000-100,000",
    raf_ipago2 == "TSI 003" ~ "<18,000",
    raf_ipago2 == "TSI 004" ~ "18,000-100,000",
    raf_ipago2 == "TSI 005" ~ ">100,000",
    T ~ "Unknown") |> factor(levels = c("Unknown", "No income", "<18,000", 
                                       "18,000-100,000", ">100,000"))) |> 
  tabyl(income)

```

# Define Comorbidities

```{r}
# Load aux bases-------------------------------------------------------------
# solve conflicts
conflicted::conflicts_prefer(lubridate::is.Date)
# Load events----------------------------------------------------------------
events <- fread2(file.path(procesados, "Target_tables_algorithm", "Big_files",
                "EVENTS.csv")) |>
  filter(person_id %in% sips_pregnancies_cohort)

# Load diagnoses-------------------------------------------------------------
 diagnoses <- fread2(file.path(procesados, "diagnoses.csv")) |>
   filter(sip %in% sips_pregnancies_cohort)

```


## Gestational diabetes 

```{r}
# Gestational diabetes-------------------------------------------------------
cod_gest_diabetes_icd9 <- c(
   "648.8" # 648.8*
   )
cod_gest_diabetes_icd10 <- c(
   "O24.4", # O24.4*
   "O99.81" # O99.81*
   )

cod_gest_diabetes <- str_c("^", 
                     c(cod_gest_diabetes_icd9, cod_gest_diabetes_icd10))

```

## Gestational Hypertension 

```{r}
# Gestational hypertension---------------------------------------------------
cod_gest_hypertension_icd9 <- c(
   "642.3" # 642.3*
   )
cod_gest_hypertension_icd10 <- c(
   "O13" # O13.** 
   )

cod_gest_hypertension <- str_c("^", 
                c(cod_gest_hypertension_icd9, cod_gest_hypertension_icd10))

```

## Preeclampsia

```{r}
# Preeclampsia---------------------------------------------------------------
cod_preeclampsia_icd9 <- c(
   "642.4", # 642.4*
   "642.5", # 642.5*
   "642.6", # 642.6*
   "642.7" # 642.7*
   )
cod_preeclampsia_icd10 <- c(
   "O11", # O11.** 
   "O14", # O14.** 
   "O15" # O15.** 
   )

cod_preeclampsia <- str_c("^", 
                c(cod_preeclampsia_icd9, cod_preeclampsia_icd10))

```

## Smoking

```{r}
# Smoking---------------------------------------------------------------
cod_smoking_icd9 <- c(
   "305.1",
   "649.0",
   "989.84",
   "V15.82"
   )
cod_smoking_icd10 <- c(
   "F17", 
   "T65.2",
   "Z72.0",
   "Z87.891",
   "O99.33"
   )

cod_smoking <- str_c("^", 
                c(cod_smoking_icd9, cod_smoking_icd10))

```

## Alcohol

```{r}
# Alcohol---------------------------------------------------------------
cod_alcohol_icd9 <- c(
"291",
"303",
"305.0",
"357.5",
"425.5",
"571.0",
"571.1",
"571.2",
"571.3",
"980.0",
"E860.0"
   )
cod_alcohol_icd10 <- c(
   "E52",
   "F10",
   "G31.2",
   "G62.1",
   "G72.1",
   "I42.6",
   "K29.2",
   "K70",
   "K85.2",
   "K86.0",
   "O35.4",
   "O99.31",
   "T51.",
   "Z71.41")

cod_alcohol <- str_c("^", 
                c(cod_alcohol_icd9, cod_alcohol_icd10))

```

## Sedentary behaviour

```{r}
# Sedentary---------------------------------------------------------------
cod_sedentary_icd9 <- c(
   "V69.0"

   )
cod_sedentary_icd10 <- c(
   "Z72.3")

cod_sedentary <- str_c("^", 
                c(cod_sedentary_icd9, cod_sedentary_icd10))

```

## Drug abuse

```{r}
# Drug-----------------------------------------------------------------------
cod_drug_abuse_icd9 <- c(
"304",
"305.2",
"305.3",
"305.4",
"305.5",
"305.6",
"305.7",
"305.8",
"305.9",
"648.30",
"655.5"
   )
cod_drug_abuse_icd10 <- c(
"F11",
"F12",
"F13",
"F14",
"F15",
"F16",
"F18",
"F19",
"O35.5",
"O99.32"
   )

cod_drug_abuse <- str_c("^", 
                c(cod_drug_abuse_icd9, cod_drug_abuse_icd10))

```

## Hypertension

```{r}
# Hypertension---------------------------------------------------------------
cod_hypert_icd9 <- c(
   "401", "402", "403", "404", "405", 
   "437.2",
   "642.0", "642.1", "642.2"
   )
cod_hypert_icd10 <- c(
    "I10", "I11", "I12", "I13", "I15", "I16", "I67.4",
    "O10"
    )

cod_hypert <- str_c("^", 
                     c(cod_hypert_icd9, cod_hypert_icd10))

```

## Congestive heart failure

```{r}
# CHF------------------------------------------------------------------------
cod_chf_icd9 <- c(
"398.91",
"402.01",
"402.11",
"402.91",
"404.01",
"404.03",
"404.11",
"404.13",
"404.91",
"404.93",
"425.4",
"428"   )

cod_chf_icd10 <- c(
"I09.81",
"I11.0",
"I13.0",
"I13.2",
"I42.0",
"I50"
    )

cod_chf <- str_c("^", 
                     c(cod_chf_icd9, cod_chf_icd10))

```

## Lipid disorders

```{r}
# lipid----------------------------------------------------------------------
cod_lipid_icd9 <- c(
   "272")
  cod_lipid_icd10 <- c(
   "E78")

cod_lipid <- str_c("^", 
                     c(cod_lipid_icd9, cod_lipid_icd10))

```

## Diabetes

```{r}
# Diabetes-------------------------------------------------------------------
cod_diabetes_icd9 <- c(
   "249",  # 249.*
   "250",  # 250.*
   "648.01", # 648.01* 
   "648.03"  # 648.03* 
   )
cod_diabetes_icd10 <- c(
   "E08", # E08.* 
   "E09", # E09.*
   "E10", # E10.*
   "E11", # E11.*
   "E13",  # E13.*
   "O24.0", # O24.0*
   "O24.1", # O24.1*
   "O24.3" # O24.3*
   )

cod_diabetes <- str_c("^", 
                     c(cod_diabetes_icd9, cod_diabetes_icd10))

```

## Depression

```{r}
# Depression-----------------------------------------------------------------
cod_depression_icd9 <- c(
  "296.2", "296.3", "298.0", 
  "300.4", "301.12", "311"
  )
cod_depression_icd10 <- c(
  "F32", "F33", "F34.1",
  "F41.8", "F53", "O90.6"
  )

cod_depression <- str_c("^", 
                     c(cod_depression_icd9, cod_depression_icd10))

```

## Anxiety

```{r}
# Anxiety--------------------------------------------------------------------
cod_anxiety_icd9 <- c(
  "293.84",
  "300.0",
  "300.2"
  )

cod_anxiety_icd10 <- c(
  "F06.4",
  "F40",
  "F41"
  )

cod_anxiety <- str_c("^", 
                     c(cod_anxiety_icd9, cod_anxiety_icd10))

```

## Psychotic disorder

```{r}
# Psychotic disorder---------------------------------------------------------
cod_psychotic_icd9 <- c(
"295",
"297",
"298",
"299")

cod_psychotic_icd10 <- c(
"F20",
"F21",
"F22",
"F23",
"F24",
"F25",
"F28",
"F29",
"F84"
  )

cod_psychotic <- str_c("^", 
                     c(cod_psychotic_icd9, cod_psychotic_icd10))

```

## Bipolar

```{r}
# Bipolar-----------------------------------------------------------------
cod_bipolar_icd9 <- c(
"296.0",
"296.1",
"296.4",
"296.5",
"296.6",
"296.7",
"296.80",
"296.89",
"296.9"
  )

cod_bipolar_icd10 <- c(
  "F30",
  "F31"
  )

cod_bipolar <- str_c("^", 
                     c(cod_bipolar_icd9, cod_bipolar_icd10))

```


## Obesity

```{r}
# Obesity--------------------------------------------------------------------
cod_obesity_icd9 <- c(
  "278.00", 
  "278.01", 
  "278.03", 
  "649.1", 
  "649.2",
  "V45.86", 
  "V85.3", 
  "V85.4"
  )
cod_obesity_icd10 <- c(
  "E66.0", 
  "E66.1", 
  "E66.2", 
  "E66.8", 
  "E66.9", 
  "O99.21", 
  "O99.84",
  "Z68.3", 
  "Z68.4",
  "Z98.84"
  )

cod_obesity <- str_c("^", 
                     c(cod_obesity_icd9, cod_obesity_icd10))

```

## Asthma

```{r}
# Asthma---------------------------------------------------------------------
cod_asthma_icd9 <- c(
  "493"
  )
cod_asthma_icd10 <- c(
  "J45"
  )

cod_asthma <- str_c("^", 
                     c(cod_asthma_icd9, cod_asthma_icd10))

```

## Epilepsy

```{r}
# Epilepsy-----------------------------------------------------------------
cod_epilepsy_icd9 <- c(
  "345"
  )
cod_epilepsy_icd10 <- c(
  "G40"
  )

cod_epilepsy <- str_c("^", 
                     c(cod_epilepsy_icd9, cod_epilepsy_icd10))

```

# Create comorbidities

```{r}
# function to search comorbidities-------------------------------------------
buscar_comor <- function (base, 
                               nombre_comor,
                               var_cod,
                               cod_comor) {
  
  base |> 
    mutate({{nombre_comor}} := if_else(if_any(starts_with(var_cod), 
                                ~str_detect(.x, paste(cod_comor,
        collapse = "|")) == TRUE), 1, 0)) 
}
  
```

## Gestational diabetes

```{r}
# create gestational diabetes comorbidity
events_gest_diabetes <- events |> 
  buscar_comor(
    nombre_comor = "m_gest_diabetes",
    var_cod = "event_code",
    cod_comor = cod_gest_diabetes) |> 
  filter(m_gest_diabetes == 1) |> 
  select(person_id, start_date_record) |> 
  group_by(person_id) |> 
  mutate(n_gest_diabetes = 1:n()) |>
  pivot_wider(
    names_from = n_gest_diabetes,
    names_prefix = "gest_diabetes_date_",
    values_from = start_date_record) |> 
  ungroup()

pregnancies_cohort_02 <- pregnancies_cohort_1b |> 
  left_join(events_gest_diabetes, by = c("sip" = "person_id")) |> 
  mutate(
    across(starts_with("gest_diabetes_date"), 
           ~ if_else(between(.x, fecha_inicio %m+% days(140), fecha_fin), 1, 0, 
                     missing = 0))) |> 
   mutate(m_gest_diabetes = pmax(
     gest_diabetes_date_1, 
     gest_diabetes_date_2, 
     gest_diabetes_date_3, 
     gest_diabetes_date_4, 
     gest_diabetes_date_5, 
     gest_diabetes_date_6, 
     gest_diabetes_date_7, 
     gest_diabetes_date_8, 
     gest_diabetes_date_9, 
     gest_diabetes_date_10, 
     gest_diabetes_date_11, 
     gest_diabetes_date_12, 
     gest_diabetes_date_13, 
     gest_diabetes_date_14, 
     gest_diabetes_date_15, 
     gest_diabetes_date_16, 
     gest_diabetes_date_17, 
     gest_diabetes_date_18, 
     gest_diabetes_date_19, 
     gest_diabetes_date_20, 
     gest_diabetes_date_21, 
     gest_diabetes_date_22, 
     gest_diabetes_date_23, 
     gest_diabetes_date_24, 
     gest_diabetes_date_25, 
     gest_diabetes_date_26, 
     gest_diabetes_date_27, 
     gest_diabetes_date_28, 
     gest_diabetes_date_29, 
     gest_diabetes_date_30, 
     gest_diabetes_date_31, 
     gest_diabetes_date_32, 
     gest_diabetes_date_33, 
     gest_diabetes_date_34, 
     gest_diabetes_date_35, 
     gest_diabetes_date_36, 
     gest_diabetes_date_37, 
     gest_diabetes_date_38, 
     gest_diabetes_date_39, 
     gest_diabetes_date_40, 
     gest_diabetes_date_41, 
     gest_diabetes_date_42, 
     gest_diabetes_date_43, 
     gest_diabetes_date_44, 
     gest_diabetes_date_45, 
     gest_diabetes_date_46, 
     gest_diabetes_date_47, 
     gest_diabetes_date_48, 
     gest_diabetes_date_49, 
     gest_diabetes_date_50, 
     gest_diabetes_date_51, 
     gest_diabetes_date_52, 
     gest_diabetes_date_53, 
     gest_diabetes_date_54, 
     gest_diabetes_date_55, 
     gest_diabetes_date_56, 
     gest_diabetes_date_57, 
     gest_diabetes_date_58, 
     gest_diabetes_date_59, 
     gest_diabetes_date_60, 
     gest_diabetes_date_61, 
     gest_diabetes_date_62, 
     gest_diabetes_date_63, 
     gest_diabetes_date_64, 
     gest_diabetes_date_65, 
                                 na.rm = TRUE)) |> 
  select(-c(gest_diabetes_date_1:gest_diabetes_date_65))
  
pregnancies_cohort_02 |> 
   tabyl(m_gest_diabetes)

```

## Gestational hypertension

```{r}
# create gestational diabetes comorbidity
events_gest_ht <- events |> 
  buscar_comor(
    nombre_comor = "m_gest_hypertension",
    var_cod = "event_code",
    cod_comor = cod_gest_hypertension) |> 
  filter(m_gest_hypertension == 1) |> 
  select(person_id, start_date_record) |> 
  group_by(person_id) |> 
  mutate(n_gest_hypertension = 1:n()) |>
  pivot_wider(
    names_from = n_gest_hypertension,
    names_prefix = "gest_hypertension_date_",
    values_from = start_date_record) |> 
  ungroup()

pregnancies_cohort_03 <- pregnancies_cohort_02 |> 
  left_join(events_gest_ht, by = c("sip" = "person_id")) |> 
  mutate(
    across(starts_with("gest_hypertension_date_"), 
           ~ if_else(between(.x, fecha_inicio %m+% days(140), fecha_fin), 1, 0, 
                     missing = 0))) |> 
   mutate(m_gest_hypertension = pmax(
     gest_hypertension_date_1, 
     gest_hypertension_date_2, 
     gest_hypertension_date_3, 
     gest_hypertension_date_4, 
     gest_hypertension_date_5, 
     gest_hypertension_date_6, 
     gest_hypertension_date_7, 
     gest_hypertension_date_8, 
     gest_hypertension_date_9, 
     gest_hypertension_date_10, 
     gest_hypertension_date_11, 
     gest_hypertension_date_12, 
     gest_hypertension_date_13, 
     gest_hypertension_date_14, 
     gest_hypertension_date_15, 
     gest_hypertension_date_16, 
     gest_hypertension_date_17, 
     gest_hypertension_date_18, 
     gest_hypertension_date_19, 
     gest_hypertension_date_20, 
     gest_hypertension_date_21, 
     gest_hypertension_date_22, 
     gest_hypertension_date_23, 
     gest_hypertension_date_24, 
     gest_hypertension_date_25, 
     gest_hypertension_date_26, 
     gest_hypertension_date_27, 
     gest_hypertension_date_28, 
     gest_hypertension_date_29, 
     gest_hypertension_date_30, 
     gest_hypertension_date_31, 
     gest_hypertension_date_32, 
     gest_hypertension_date_33, 
     gest_hypertension_date_34, 
     gest_hypertension_date_35, 
     gest_hypertension_date_36, 
     gest_hypertension_date_37, 
     gest_hypertension_date_38, 
     gest_hypertension_date_39, 
     gest_hypertension_date_40, 
     gest_hypertension_date_41, 
     gest_hypertension_date_42, 
     gest_hypertension_date_43, 
     gest_hypertension_date_44, 
     gest_hypertension_date_45, 
     gest_hypertension_date_46, 
     gest_hypertension_date_47, 
     gest_hypertension_date_48,
                                 na.rm = TRUE)) |> 
  select(-c(gest_hypertension_date_1:gest_hypertension_date_48))
  
pregnancies_cohort_03 |> 
   tabyl(m_gest_hypertension)

```

## Preeclampsia

```{r}
# create gestational diabetes comorbidity
events_preeclampsia <- events |> 
  buscar_comor(
    nombre_comor = "m_preeclampsia",
    var_cod = "event_code",
    cod_comor = cod_preeclampsia) |> 
  filter(m_preeclampsia == 1) |> 
  select(person_id, start_date_record) |> 
  group_by(person_id) |> 
  mutate(n_preeclampsia = 1:n()) |>
  pivot_wider(
    names_from = n_preeclampsia,
    names_prefix = "preeclampsia_date_",
    values_from = start_date_record) |> 
  ungroup()

pregnancies_cohort_04 <- pregnancies_cohort_03 |> 
  left_join(events_preeclampsia, by = c("sip" = "person_id")) |> 
  mutate(
    across(starts_with("preeclampsia_date_"), 
           ~ if_else(between(.x, fecha_inicio %m+% days(140), fecha_fin), 1, 0, 
                     missing = 0))) |> 
   mutate(m_preeclampsia = pmax(
     preeclampsia_date_1, 
     preeclampsia_date_2, 
     preeclampsia_date_3, 
     preeclampsia_date_4, 
     preeclampsia_date_5, 
     preeclampsia_date_6, 
     preeclampsia_date_7, 
     preeclampsia_date_8, 
     preeclampsia_date_9, 
     preeclampsia_date_10, 
     preeclampsia_date_11, 
     preeclampsia_date_12, 
     preeclampsia_date_13, 
     preeclampsia_date_14, 
     preeclampsia_date_15, 
     preeclampsia_date_16, 
     preeclampsia_date_17, 
     preeclampsia_date_18, 
     preeclampsia_date_19, 
     preeclampsia_date_20, 
     preeclampsia_date_21, 
     preeclampsia_date_22, 
     preeclampsia_date_23, 
     preeclampsia_date_24, 
     preeclampsia_date_25, 
     preeclampsia_date_26, 
     preeclampsia_date_27, 
     preeclampsia_date_28, 
     preeclampsia_date_29, 
     preeclampsia_date_30, 
     preeclampsia_date_31, 
     preeclampsia_date_32, 
     preeclampsia_date_33, 
     preeclampsia_date_34, 
     preeclampsia_date_35, 
     preeclampsia_date_36, 
     preeclampsia_date_37,
                                 na.rm = TRUE)) |> 
  select(-c(preeclampsia_date_1:preeclampsia_date_37))
  
pregnancies_cohort_04 |> 
   tabyl(m_preeclampsia)

```

## Smoking

```{r}
# retrieve codes-------------------------------------------------------------
events_smoking <- events |> 
  buscar_comor(
    nombre_comor = "m_smoking",
    var_cod = "event_code",
    cod_comor = cod_smoking) |> 
  filter(m_smoking == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_smoking <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_smoking",
    var_cod = "diag_cod",
    cod_comor = cod_smoking) |> 
  filter(m_smoking == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_smoking <- pregnancies_cohort_04 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_smoking, by = c("sip" = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_smoking = 1)

pregnancies_cohort_diagnoses_smoking <- pregnancies_cohort_04 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_smoking, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_smoking = 1)

pregnancies_cohort_smoking <- pregnancies_cohort_events_smoking |> 
  union(pregnancies_cohort_diagnoses_smoking)

pregnancies_cohort_05 <- pregnancies_cohort_04 |> 
  left_join(pregnancies_cohort_smoking) |> 
  mutate(m_smoking = if_else(is.na(m_smoking), 0, m_smoking))

pregnancies_cohort_05 |> tabyl(m_smoking)

```

## Alcohol

```{r}
# retrieve codes-------------------------------------------------------------
events_alcohol <- events |> 
  buscar_comor(
    nombre_comor = "m_alcohol",
    var_cod = "event_code",
    cod_comor = cod_alcohol) |> 
  filter(m_alcohol == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_alcohol <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_alcohol",
    var_cod = "diag_cod",
    cod_comor = cod_alcohol) |> 
  filter(m_alcohol == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_alcohol <- pregnancies_cohort_05 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_alcohol, by = c("sip" = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_alcohol = 1)

pregnancies_cohort_diagnoses_alcohol <- pregnancies_cohort_05 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_alcohol, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_alcohol = 1)

pregnancies_cohort_alcohol <- pregnancies_cohort_events_alcohol |> 
  union(pregnancies_cohort_diagnoses_alcohol)

pregnancies_cohort_06 <- pregnancies_cohort_05 |> 
  left_join(pregnancies_cohort_alcohol) |> 
  mutate(m_alcohol = if_else(is.na(m_alcohol), 0, m_alcohol))

pregnancies_cohort_06 |> tabyl(m_alcohol)

```

## Sedentary behaviour

```{r}
# retrieve codes-------------------------------------------------------------
events_sedentary <- events |> 
  buscar_comor(
    nombre_comor = "m_sedentary",
    var_cod = "event_code",
    cod_comor = cod_sedentary) |> 
  filter(m_sedentary == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_sedentary <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_sedentary",
    var_cod = "diag_cod",
    cod_comor = cod_sedentary) |> 
  filter(m_sedentary == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_sedentary <- pregnancies_cohort_06 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_sedentary, by = c("sip" = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_sedentary = 1)

pregnancies_cohort_diagnoses_sedentary <- pregnancies_cohort_06 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_sedentary, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_sedentary = 1)

pregnancies_cohort_sedentary <- pregnancies_cohort_events_sedentary |> 
  union(pregnancies_cohort_diagnoses_sedentary)

pregnancies_cohort_07 <- pregnancies_cohort_06 |> 
  left_join(pregnancies_cohort_sedentary) |> 
  mutate(m_sedentary = if_else(is.na(m_sedentary), 0, m_sedentary))

pregnancies_cohort_07 |> tabyl(m_sedentary)

```

## Drug abuse

```{r}
# retrieve codes-------------------------------------------------------------
events_drug_abuse <- events |> 
  buscar_comor(
    nombre_comor = "m_drug_abuse",
    var_cod = "event_code",
    cod_comor = cod_drug_abuse) |> 
  filter(m_drug_abuse == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_drug_abuse <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_drug_abuse",
    var_cod = "diag_cod",
    cod_comor = cod_drug_abuse) |> 
  filter(m_drug_abuse == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_drug_abuse <- pregnancies_cohort_07 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_drug_abuse, by = c("sip" = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_drug_abuse = 1)

pregnancies_cohort_diagnoses_drug_abuse <- pregnancies_cohort_07 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_drug_abuse, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_drug_abuse = 1)

pregnancies_cohort_drug_abuse <- pregnancies_cohort_events_drug_abuse |> 
  union(pregnancies_cohort_diagnoses_drug_abuse)

pregnancies_cohort_08 <- pregnancies_cohort_07 |> 
  left_join(pregnancies_cohort_drug_abuse) |> 
  mutate(m_drug_abuse = if_else(is.na(m_drug_abuse), 0, m_drug_abuse))

pregnancies_cohort_08 |> tabyl(m_drug_abuse)

```

## Hypertension

```{r}
# retrieve codes-------------------------------------------------------------
events_ht <- events |> 
  buscar_comor(
    nombre_comor = "m_hypertension",
    var_cod = "event_code",
    cod_comor = cod_hypert) |> 
  filter(m_hypertension == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_ht <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_hypertension",
    var_cod = "diag_cod",
    cod_comor = cod_hypert) |> 
  filter(m_hypertension == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_ht <- pregnancies_cohort_08 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_ht, by = c("sip" = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio %m+% days(140)) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_hypertension = 1)

pregnancies_cohort_diagnoses_ht <- pregnancies_cohort_08 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_ht, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio %m+% days(140)) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_hypertension = 1)

pregnancies_cohort_ht <- pregnancies_cohort_events_ht |> 
  union(pregnancies_cohort_diagnoses_ht)

pregnancies_cohort_09 <- pregnancies_cohort_08 |> 
  left_join(pregnancies_cohort_ht) |> 
  mutate(m_hypertension = if_else(is.na(m_hypertension), 0, m_hypertension))

pregnancies_cohort_09 |> tabyl(m_hypertension)

```

## Congestive heart failure

```{r}
# retrieve codes-------------------------------------------------------------
events_chf <- events |> 
  buscar_comor(
    nombre_comor = "m_chf",
    var_cod = "event_code",
    cod_comor = cod_chf) |> 
  filter(m_chf == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_chf <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_chf",
    var_cod = "diag_cod",
    cod_comor = cod_chf) |> 
  filter(m_chf == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_chf <- pregnancies_cohort_09 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_chf, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_chf = 1)

pregnancies_cohort_diagnoses_chf <- pregnancies_cohort_09 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_chf, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_chf = 1)

pregnancies_cohort_chf <- pregnancies_cohort_events_chf |> 
  union(pregnancies_cohort_diagnoses_chf)

pregnancies_cohort_10 <- pregnancies_cohort_09 |> 
  left_join(pregnancies_cohort_chf) |> 
  mutate(m_chf = if_else(is.na(m_chf), 0, m_chf))

pregnancies_cohort_10 |> tabyl(m_chf)

```

## Lipid disorders

```{r}
# retrieve codes-------------------------------------------------------------
events_lipid <- events |> 
  buscar_comor(
    nombre_comor = "m_lipid",
    var_cod = "event_code",
    cod_comor = cod_lipid) |> 
  filter(m_lipid == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_lipid <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_lipid",
    var_cod = "diag_cod",
    cod_comor = cod_lipid) |> 
  filter(m_lipid == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_lipid <- pregnancies_cohort_10 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_lipid, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_lipid = 1)

pregnancies_cohort_diagnoses_lipid <- pregnancies_cohort_10 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_lipid, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_lipid = 1)

pregnancies_cohort_lipid <- pregnancies_cohort_events_lipid |> 
  union(pregnancies_cohort_diagnoses_lipid)

pregnancies_cohort_11 <- pregnancies_cohort_10 |> 
  left_join(pregnancies_cohort_lipid) |> 
  mutate(m_lipid = if_else(is.na(m_lipid), 0, m_lipid))

pregnancies_cohort_11 |> tabyl(m_lipid)

```

## Diabetes

```{r}
# retrieve codes-------------------------------------------------------------
events_diabetes <- events |> 
  buscar_comor(
    nombre_comor = "m_diabetes",
    var_cod = "event_code",
    cod_comor = cod_diabetes) |> 
  filter(m_diabetes == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_diabetes <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_diabetes",
    var_cod = "diag_cod",
    cod_comor = cod_diabetes) |> 
  filter(m_diabetes == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_diabetes <- pregnancies_cohort_11 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_diabetes, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio %m+% days(140)) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_diabetes = 1)

pregnancies_cohort_diagnoses_diabetes <- pregnancies_cohort_11 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_diabetes, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio %m+% days(140)) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_diabetes = 1)

pregnancies_cohort_diabetes <- pregnancies_cohort_events_diabetes |> 
  union(pregnancies_cohort_diagnoses_diabetes)

pregnancies_cohort_12 <- pregnancies_cohort_11 |> 
  left_join(pregnancies_cohort_diabetes) |> 
  mutate(m_diabetes = if_else(is.na(m_diabetes), 0, m_diabetes))

pregnancies_cohort_12 |> tabyl(m_diabetes)

```

## Depression

```{r}
# retrieve codes-------------------------------------------------------------
events_depression <- events |> 
  buscar_comor(
    nombre_comor = "m_depression",
    var_cod = "event_code",
    cod_comor = cod_depression) |> 
  filter(m_depression == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_depression <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_depression",
    var_cod = "diag_cod",
    cod_comor = cod_depression) |> 
  filter(m_depression == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_depression <- pregnancies_cohort_12 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_depression, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_depression = 1)

pregnancies_cohort_diagnoses_depression <- pregnancies_cohort_12 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_depression, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_depression = 1)

pregnancies_cohort_depression <- pregnancies_cohort_events_depression |> 
  union(pregnancies_cohort_diagnoses_depression)

pregnancies_cohort_13 <- pregnancies_cohort_12 |> 
  left_join(pregnancies_cohort_depression) |> 
  mutate(m_depression = if_else(is.na(m_depression), 0, m_depression))

pregnancies_cohort_13 |> tabyl(m_depression)

```

## Anxiety

```{r}
# retrieve codes-------------------------------------------------------------
events_anxiety <- events |> 
  buscar_comor(
    nombre_comor = "m_anxiety",
    var_cod = "event_code",
    cod_comor = cod_anxiety) |> 
  filter(m_anxiety == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_anxiety <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_anxiety",
    var_cod = "diag_cod",
    cod_comor = cod_anxiety) |> 
  filter(m_anxiety == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_anxiety <- pregnancies_cohort_13 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_anxiety, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_anxiety = 1)

pregnancies_cohort_diagnoses_anxiety <- pregnancies_cohort_13 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_anxiety, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_anxiety = 1)

pregnancies_cohort_anxiety <- pregnancies_cohort_events_anxiety |> 
  union(pregnancies_cohort_diagnoses_anxiety)

pregnancies_cohort_14 <- pregnancies_cohort_13 |> 
  left_join(pregnancies_cohort_anxiety) |> 
  mutate(m_anxiety = if_else(is.na(m_anxiety), 0, m_anxiety))

pregnancies_cohort_14 |> tabyl(m_anxiety)

```

## Psychotic disorder

```{r}
# retrieve codes-------------------------------------------------------------
events_psychotic <- events |> 
  buscar_comor(
    nombre_comor = "m_psychotic",
    var_cod = "event_code",
    cod_comor = cod_psychotic) |> 
  filter(m_psychotic == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_psychotic <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_psychotic",
    var_cod = "diag_cod",
    cod_comor = cod_psychotic) |> 
  filter(m_psychotic == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_psychotic <- pregnancies_cohort_14 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_psychotic, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_psychotic = 1)

pregnancies_cohort_diagnoses_psychotic <- pregnancies_cohort_14 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_psychotic, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_psychotic = 1)

pregnancies_cohort_psychotic <- pregnancies_cohort_events_psychotic |> 
  union(pregnancies_cohort_diagnoses_psychotic)

pregnancies_cohort_15 <- pregnancies_cohort_14 |> 
  left_join(pregnancies_cohort_psychotic) |> 
  mutate(m_psychotic = if_else(is.na(m_psychotic), 0, m_psychotic))

pregnancies_cohort_15 |> tabyl(m_psychotic)

```

## Bipolar

```{r}
# retrieve codes-------------------------------------------------------------
events_bipolar <- events |> 
  buscar_comor(
    nombre_comor = "m_bipolar",
    var_cod = "event_code",
    cod_comor = cod_bipolar) |> 
  filter(m_bipolar == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_bipolar <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_bipolar",
    var_cod = "diag_cod",
    cod_comor = cod_bipolar) |> 
  filter(m_bipolar == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_bipolar <- pregnancies_cohort_15 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_bipolar, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_bipolar = 1)

pregnancies_cohort_diagnoses_bipolar <- pregnancies_cohort_15 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_bipolar, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_bipolar = 1)

pregnancies_cohort_bipolar <- pregnancies_cohort_events_bipolar |> 
  union(pregnancies_cohort_diagnoses_bipolar)

pregnancies_cohort_16 <- pregnancies_cohort_15 |> 
  left_join(pregnancies_cohort_bipolar) |> 
  mutate(m_bipolar = if_else(is.na(m_bipolar), 0, m_bipolar))

pregnancies_cohort_16 |> tabyl(m_bipolar)

```

## Obesity

```{r}
# retrieve codes-------------------------------------------------------------
events_obesity <- events |> 
  buscar_comor(
    nombre_comor = "m_obesity",
    var_cod = "event_code",
    cod_comor = cod_obesity) |> 
  filter(m_obesity == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_obesity <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_obesity",
    var_cod = "diag_cod",
    cod_comor = cod_obesity) |> 
  filter(m_obesity == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_obesity <- pregnancies_cohort_16 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_obesity, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_obesity = 1)

pregnancies_cohort_diagnoses_obesity <- pregnancies_cohort_16 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_obesity, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_obesity = 1)

pregnancies_cohort_obesity <- pregnancies_cohort_events_obesity |> 
  union(pregnancies_cohort_diagnoses_obesity)

pregnancies_cohort_17 <- pregnancies_cohort_16 |> 
  left_join(pregnancies_cohort_obesity) |> 
  mutate(m_obesity = if_else(is.na(m_obesity), 0, m_obesity))

pregnancies_cohort_17 |> tabyl(m_obesity)

```

## Asthma

```{r}
# retrieve codes-------------------------------------------------------------
events_asthma <- events |> 
  buscar_comor(
    nombre_comor = "m_asthma",
    var_cod = "event_code",
    cod_comor = cod_asthma) |> 
  filter(m_asthma == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_asthma <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_asthma",
    var_cod = "diag_cod",
    cod_comor = cod_asthma) |> 
  filter(m_asthma == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_asthma <- pregnancies_cohort_17 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_asthma, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_asthma = 1)

pregnancies_cohort_diagnoses_asthma <- pregnancies_cohort_17 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_asthma, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_asthma = 1)

pregnancies_cohort_asthma <- pregnancies_cohort_events_asthma |> 
  union(pregnancies_cohort_diagnoses_asthma)

pregnancies_cohort_18 <- pregnancies_cohort_17 |> 
  left_join(pregnancies_cohort_asthma) |> 
  mutate(m_asthma = if_else(is.na(m_asthma), 0, m_asthma))

pregnancies_cohort_18 |> tabyl(m_asthma)

```

## Epilepsy

```{r}
# retrieve codes-------------------------------------------------------------
events_epilepsy <- events |> 
  buscar_comor(
    nombre_comor = "m_epilepsy",
    var_cod = "event_code",
    cod_comor = cod_epilepsy) |> 
  filter(m_epilepsy == 1) |> 
  select(person_id, start_date_record) |> 
  tidylog::distinct() 

diagnoses_epilepsy <- diagnoses |> 
  buscar_comor(
    nombre_comor = "m_epilepsy",
    var_cod = "diag_cod",
    cod_comor = cod_epilepsy) |> 
  filter(m_epilepsy == 1) |> 
  select(sip, fecha_act, fecha_desact) |> 
  tidylog::distinct() 

pregnancies_cohort_events_epilepsy <- pregnancies_cohort_18 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(events_epilepsy, by = c(sip = "person_id")) |>  
   tidylog::filter(
     start_date_record >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    start_date_record <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_epilepsy = 1)

pregnancies_cohort_diagnoses_epilepsy <- pregnancies_cohort_18 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(diagnoses_epilepsy, by = c("sip" = "sip")) |>  
   tidylog::filter(
     is.na(fecha_desact) |
     fecha_desact >= fecha_inicio %m-% days(180)) |> 
  tidylog::filter(
    fecha_act <= fecha_inicio) |> 
  tidylog::distinct(sip, fecha_inicio) |> 
  mutate(m_epilepsy = 1)

pregnancies_cohort_epilepsy <- pregnancies_cohort_events_epilepsy |> 
  union(pregnancies_cohort_diagnoses_epilepsy)

pregnancies_cohort_19 <- pregnancies_cohort_18 |> 
  left_join(pregnancies_cohort_epilepsy) |> 
  mutate(m_epilepsy = if_else(is.na(m_epilepsy), 0, m_epilepsy))

pregnancies_cohort_19 |> tabyl(m_epilepsy)

```

# Visits

```{r}
# # Load visits----------------------------------------------------------------
#  visit_occurrence <- fread2(file.path(procesados, 
#                  "Target_tables_algorithm", "Big_files",
#                  "VISIT_OCCURRENCE.csv")) |>
#    filter(person_id %in% sips_pregnancies_cohort)
# 
# # read pcv file--------------------------------------------------------------
# fread_pcv <- function(...){
#   data.table::fread(
#     encoding = "UTF-8",
#     colClasses = c(
#     sip                  = "character",
#     fecha_consulta       = "Date",
#     serv_at_cod          = "character",
#     serv_at_desc         = "character",
#     diag_cod             = "character",
#     diag_desc            = "character",
#     tipo_codigo          = "character"),
#     ...)
# } 
# 
# pcv <- fread_pcv(file.path(procesados, "pcv.csv")) |>
#    filter(sip %in% sips_pregnancies_cohort)
# 
# # read cex file--------------------------------------------------------------
# fread_cex <- function(...){
#   data.table::fread(
#     encoding = "UTF-8",
#     colClasses = c(
#     sip                = "character",
#     fecha_consulta     = "Date",
#     especialidad_cod   = "character",
#     especialidad_desc  = "character",
#     d1_cod             = "character",
#     d1_desc            = "character",
#     d2_cod             = "character",
#     d2_desc            = "character",
#     d3_cod             = "character",
#     d3_desc            = "character",
#     d4_cod             = "character",
#     d4_desc            = "character",
#     tipo_codigo1       = "character",
#     tipo_codigo2       = "character",
#     tipo_codigo3       = "character",
#     tipo_codigo4       = "character"),
#     ...)
# }
# 
# cex <- fread_pcv(file.path(procesados, "cex.csv")) |>
#    filter(sip %in% sips_pregnancies_cohort)

```

## use of services

```{r}
# # use of services------------------------------------------------------------
# visit_occurrence_cohort <- visit_occurrence |> 
#   tidylog::filter(origin_of_visit != "DIAGNOSES")
# 
# visit_occurrence_cohort |> count(meaning_of_visit)
# 
# pregnancies_cohort_visits <- pregnancies_cohort_08 |> 
# # unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
#   left_join(visit_occurrence_cohort, by = "person_id") |>  
#    tidylog::filter(visit_start_date >= pregnancy_start_date) |> 
#    tidylog::filter(visit_start_date <= pregnancy_end_date) 
# 
# pregnancies_cohort_visits2 <- pregnancies_cohort_visits |> 
#   group_by(pregnancy_id) |> 
#   summarise(
#     n_emergency_contact = sum(meaning_of_visit == "emergency_contact"),
#     n_hospitalisation = sum(meaning_of_visit == "hospitalisation"),
#     n_hospitalisation_not_overnight = sum(meaning_of_visit == "hospitalisation_not_overnight"),
#     n_oupatient_specialist_visit = sum(meaning_of_visit == "oupatient_specialist_visit"),
#     n_primary_care = sum(meaning_of_visit == "primary_care"))
# 
# pregnancies_cohort_09 <- pregnancies_cohort_08 |> 
#   left_join(pregnancies_cohort_visits2) |> 
#   mutate(across(n_emergency_contact:n_primary_care, 
#                 ~if_else(is.na(.x), 0, .x)))


```

# Outcomes

## Cong

```{r}
# load cong------------------------------------------------------------------
cong <- fread2(file.path(procesados, "cong.csv"))

```

```{r}
pregnancies_cohort_cong_0 <- cong |> 
  select(sip, fecha_fin_emb) 

pregnancies_cohort_cong <- pregnancies_cohort_19 |> 
  left_join(pregnancies_cohort_cong_0, by = c("sip" = "sip")) |> 
  tidylog::filter(fecha_fin_emb == fecha_fin) |> 
  tidylog::distinct(sip, fecha_fin) |> 
  mutate(o_cong = 1)  

pregnancies_cohort_20 <- pregnancies_cohort_19 |> 
  left_join(pregnancies_cohort_cong) |> 
  mutate(o_cong = if_else(is.na(o_cong), 0, o_cong))

pregnancies_cohort_20 |> tabyl(o_cong)

```

## Preterm

```{r}
# make preterm
pregnancies_cohort_21 <- pregnancies_cohort_20 |> 
  mutate(o_preterm = if_else(duration < 259, 1, 0))
  
pregnancies_cohort_21 |> filter(tipo_fin == "livebirth") |> 
  tabyl(o_preterm)

```

## Low weight at birth

```{r}
# read pregnancy_indicators--------------------------------------------------
mdr <- fread2(file.path(procesados, "mdr.csv"))

pregnancies_cohort_peso <- mdr |> 
  filter(
  year(fecha_fin_emb) >= 2009,
  year(fecha_fin_emb) <= 2021) |> 
  mutate(bajo_peso = if_else(peso < 2500, 1, 0))

pregnancies_cohort_peso |> tabyl(bajo_peso)

saveRDS(pregnancies_cohort_peso, file.path(procesados, "pregnancy_algorithm",
                                          "pregnancies_cohort_peso.RDS"))

mdr |> summarise(
  mean_peso = mean(peso, na.rm = TRUE),
  sd_peso = sd(peso, na.rm = TRUE)
)

```

```{r}
# agrupar partos en embarazos------------------------------------------------
pregnancies_cohort_peso2 <- pregnancies_cohort_peso |> 
  select(sip = sip_madre, 
         fecha_fin = fecha_fin_emb,
         bajo_peso) |> 
  group_by(sip, fecha_fin) |>
  summarise(bajo_peso = max(bajo_peso))

pregnancies_cohort_22 <- pregnancies_cohort_21 |> 
  left_join(pregnancies_cohort_peso2, by = c("sip", 
                                             "fecha_fin"))

```


```{r}
# free some memory-----------------------------------------------------------
rm(
  pregnacies_cohort,
  pregnancies_cohort_0,
  pregnancies_cohort_02,
  pregnancies_cohort_03,
  pregnancies_cohort_04,
  pregnancies_cohort_05,
  pregnancies_cohort_06,
  pregnancies_cohort_07,
  pregnancies_cohort_08,
  pregnancies_cohort_09,
  pregnancies_cohort_0b,
  pregnancies_cohort_1,
  pregnancies_cohort_10,
  pregnancies_cohort_11,
  pregnancies_cohort_12,
  pregnancies_cohort_13,
  pregnancies_cohort_14,
  pregnancies_cohort_15,
  pregnancies_cohort_16,
  pregnancies_cohort_17,
  pregnancies_cohort_18,
  pregnancies_cohort_19,
  pregnancies_cohort_20
)

gc()

```

# C-Section

```{r}
# c-section codes--------------------------------------------------------------
c_section_codes <- c(
  "^O82",
  "^O82.0",
  "^O82.1",
  "^O82.2",
  "^O82.8",
  "^O82.9",
  "^O84.2",
  "^O75.82",
  "^Z38.01",
  "^Z38.31",
  "^Z38.62",
  "^Z38.64",
  "^Z38.66",
  "^Z38.69",
  "^669.7")

proc_c_section <- c("^74$",
  "^74.9",
  "^74.91",
  "^74.99",
  "^74.0",
  "^74.1",
  "^74.2",
  "^74.3",
  "^74.4", 
  "10D00Z0", "10D00Z1", "10D00Z2")

```

```{r}
# load tables-----------------------------------------------------------------
procedures <- fread2(file.path(procesados, "Target_tables_algorithm",
                "procedures.csv")) |>
  filter(person_id %in% sips_pregnancies_cohort)

```

```{r}
# count person with cesarean codes-------------------------------------------
events_c_section <- events |> 
  mutate(c_section = if_else(str_detect(event_code, 
         paste(c_section_codes, collapse = "|")) == TRUE, 1, 0)) |> 
  # seleccionar nicamente los partos con cesrea
  filter(c_section == 1) |> 
  mutate(record_date = ymd(start_date_record)) |> 
  filter(record_date >= ymd(20090401)) |> 
  filter(record_date < ymd(20200101)) |> 
  distinct(person_id, record_date) 

procedures_c_section <- procedures |>
  mutate(c_section = if_else(str_detect(procedure_code, 
         paste0(proc_c_section, collapse="|")) == TRUE, 1, 0)) |> 
  # seleccionar nicamente los partos con cesrea
  filter(c_section == 1) |> 
 mutate(record_date = ymd(procedure_date)) |> 
  filter(record_date >= ymd(20090401)) |> 
  filter(record_date < ymd(20211231)) |> 
  distinct(person_id, record_date) 

events_c_section |> union(procedures_c_section) |> 
  distinct(person_id, year(record_date))

# check number of c-section  
numerator_cesarean <- events_c_section |> union(procedures_c_section) |> 
  tidylog::distinct(person_id, year(record_date)) |> nrow()
denominator_cesarean <- pregnancies_cohort_22 |> 
  filter(tipo_fin == "livebirth") |> nrow()

percent_cesarean <- numerator_cesarean/denominator_cesarean

```


```{r}
# append c-section to the table----------------------------------------------
c_sections_with_date <- events_c_section |> 
  union(procedures_c_section) |> 
  distinct(person_id, record_date)


pregnancies_cohort_c_section <- pregnancies_cohort_22 |> 
# unir cmbd con cohorte y filtrar fechas que no coincidan con el follow-up
  left_join(c_sections_with_date, by = c(sip = "person_id")) |>  
   tidylog::filter(
     record_date >= fecha_fin %m-% days(14)) |> 
  tidylog::filter(
    record_date <= fecha_fin %m+% days(14)) |> 
  tidylog::distinct(sip, fecha_fin) |> 
  mutate(c_section = 1)

pregnancies_cohort_23 <- pregnancies_cohort_22 |> 
  left_join(pregnancies_cohort_c_section) |> 
  mutate(c_section = if_else(is.na(c_section) | tipo_fin != "livebirth",
                             0, c_section))

pregnancies_cohort_23 |> 
  tidylog::filter(tipo_fin == "livebirth") |>
  tabyl(tipo_fin, c_section)

```

# Drug use

```{r}
# load disp_clean------------------------------------------------------------
gaia_pq <- open_dataset(file.path(procesados, "07_GAIA", 
                                  "07j_gaia_non_standard"))

gaia <- gaia_pq |> collect()

```

## Prescription

```{r}
# create gaia-preg-----------------------------------------------------------
pregnancies_cohort_pres <- pregnancies_cohort_23 |> 
  left_join(gaia, by = "sip") |>  
  tidylog::filter(fecha_receta >= fecha_inicio %m-% weeks(13), 
         fecha_receta <= fecha_fin %m+% weeks(13)) |> 
  mutate(
    # 13wks before
    fecha_inicio_13s_antes = fecha_inicio %m-% weeks(13),
    fecha_fin_13s_antes = fecha_inicio %m-% days(1),
    # first trim
    fecha_inicio_1t = fecha_inicio,
    fecha_fin_1t = if_else(
      fecha_fin >= fecha_inicio %m+% days(97),
      fecha_inicio %m+% days(97), fecha_fin),
    # second trim     
    fecha_inicio_2t = if_else(
    fecha_fin >= fecha_inicio %m+% days(98),
      fecha_inicio %m+% days(98), NA_Date_),
    fecha_fin_2t = if_else(
      fecha_fin >= fecha_inicio %m+% days(98),
      pmin(fecha_inicio %m+% days(195), fecha_fin), NA_Date_),
    # third trim
    fecha_inicio_3t = if_else(
      fecha_fin >= fecha_inicio %m+% days(196),
      fecha_inicio %m+% days(196), NA_Date_),
    fecha_fin_3t = if_else(
      fecha_fin >= fecha_inicio %m+% days(196),
      fecha_fin, NA_Date_),
    # 13wks after
    fecha_inicio_13s_despues = fecha_fin %m+% days(1),
    fecha_fin_13s_despues = fecha_fin %m+% weeks(13)
    ) |>
  mutate(periodo_exposicion = case_when(
    between(fecha_receta, fecha_inicio_13s_antes, fecha_fin_13s_antes) ~ 0,
    between(fecha_receta, fecha_inicio_1t, fecha_fin_1t) ~ 1,
    between(fecha_receta, fecha_inicio_2t, fecha_fin_2t) ~ 2,
    between(fecha_receta, fecha_inicio_3t, fecha_fin_3t) ~ 3,
    between(fecha_receta, fecha_inicio_13s_despues, fecha_fin_13s_despues) ~ 4))

pregnancies_cohort_pres |> tabyl(periodo_exposicion)

```


```{r}
# at least one prescription--------------------------------------------------
pregnancies_cohort_pres |>
  tidylog::filter(periodo_exposicion %in% 1:3) |> 
  tidylog::distinct(sip, fecha_inicio)

pregnancies_cohort_pres2 <- pregnancies_cohort_pres |>
  tidylog::filter(periodo_exposicion %in% 1:3) |> 
  group_by(sip, fecha_inicio) |> 
  summarise(n_prin_act_cod_pres = n_distinct(prin_act_cod))

pregnancies_cohort_24 <- pregnancies_cohort_23 |>
  left_join(pregnancies_cohort_pres2) |> 
  mutate(n_prin_act_cod_pres = if_else(is.na(n_prin_act_cod_pres), 
                                       0, n_prin_act_cod_pres)) |> 
  mutate(
  at_least_one_drug_pres = if_else(n_prin_act_cod_pres > 0, 1, 0))

pregnancies_cohort_24 |> tabyl(at_least_one_drug_pres)

```

## Dispensing

```{r}
# at least one prescription--------------------------------------------------
pregnancies_cohort_disp <- pregnancies_cohort_pres |>
  tidylog::filter(periodo_exposicion %in% 1:3) |> 
  tidylog::filter(tipo_receta != "prescription") |> 
  group_by(sip, fecha_inicio) |> 
  summarise(n_prin_act_cod_disp = n_distinct(prin_act_cod))

pregnancies_cohort_25 <- pregnancies_cohort_24 |>
  left_join(pregnancies_cohort_disp) |> 
  mutate(n_prin_act_cod_disp = if_else(is.na(n_prin_act_cod_disp), 
                                       0, n_prin_act_cod_disp)) |> 
  mutate(
  at_least_one_drug_disp = if_else(n_prin_act_cod_disp > 0, 1, 0))

pregnancies_cohort_25 |> tabyl(at_least_one_drug_disp)

pregnancies_cohort_25 |> summarise(
  mean(n_prin_act_cod_pres),
  mean(n_prin_act_cod_disp),
  sd(n_prin_act_cod_pres),
  sd(n_prin_act_cod_disp))

```

```{r}
# add previous use of drug---------------------------------------------------
pregnancies_cohort_disp_use <- pregnancies_cohort_pres |>
  tidylog::filter(tipo_receta != "prescription") |> 
  tidylog::filter(periodo_exposicion == 0) |> 
  distinct(sip, fecha_inicio) |> 
  mutate(prev_drug_use = 1)

pregnancies_cohort_26 <- pregnancies_cohort_25 |>
  left_join(pregnancies_cohort_disp_use) |> 
  mutate(prev_drug_use = if_else(is.na(prev_drug_use), 0, prev_drug_use))

pregnancies_cohort_26 |> tabyl(prev_drug_use)

```


# Save base

```{r}
pregnancies_cohort_table_6m_lookback <- pregnancies_cohort_26

```

```{r}
# change the misclassified imputed pregnancies
pregnancies_cohort_table_6m_lookback_0 <- readRDS(file.path(procesados, 
                                  "pregnancy_algorithm",
                                  "pregnancies_cohort_table_6m_lookback.RDS"))

pregnancies_cohort_table_6m_lookback_0 |> filter(sip %in% 
                  c("VZyDjt87ZXA4+Mq8WPq2yA==", 
                    "l8Uv10Q1vEJJdZ8HYWFA9Q=="))

pregnancies_misclasified_as_imputed <- c(
# con peso
  230913,
  336913,
# preterm
  56553,
  437855,
  437916,
  438072)

pregnancies_cohort_table_6m_lookback <- pregnancies_cohort_table_6m_lookback_0 |> 
  mutate(imputed = if_else(pregnancy_id %in% pregnancies_misclasified_as_imputed,
                           "no", imputed))

# check
pregnancies_cohort_table_6m_lookback |> 
  filter(pregnancy_id %in% pregnancies_misclasified_as_imputed)

# OK

```


```{r}
saveRDS(pregnancies_cohort_table_6m_lookback, file.path(procesados, "pregnancy_algorithm",
                                  "pregnancies_cohort_table_6m_lookback.RDS"))

```

## Trimesters

- Trimester 1: from Last Menstrual Period (LMP) to day < 98 after LMP; or end of pregnancy, whichever earlier
- Trimester 2: from day 98 after LMP to day <196 after LMP; or end of pregnancy, whichever earlier
- Trimester 3: from day 196 after LMP onwards until end of pregnancy

```{r}
#| include: false
# copy .qmd to CERES---------------------------------------------------------
unlink(file.path("r:", "PREGVAL", "2-SCRIPTS", 
                 "3_43_PREGVAL_Cohort_Profile_Results_6m_lookback.qmd"))
file.copy(from = file.path(
          "3_43_PREGVAL_Cohort_Profile_Results_6m_lookback.qmd"),
          to = file.path("r:", "PREGVAL", "2-SCRIPTS"))

```

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
